---
title: "README"
author: "John Helveston"
date: "4/30/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# stanTuner
This code uses the algebra solver in [Stan](https://mc-stan.org/) to find the parameters of a distribution that produce a desired tail behavior. This can be a useful tool when choosing parameters for prior distributions. Here's how to use it:

1. Choose a distribution.
2. Define the quantile boundaries and the amount of probability density that you wish to have above and below those boundaries.
3. Let Stan go find the parameters that produce the desired distribution.

Currently supported distributions:

* Normal
* Log-Normal
* Beta
* Inverse Gamma

# Required libraries:

If you don't want to use the shiny app interface, you only need to have the `rstan` library installed (and obviously also have [Stan](https://mc-stan.org/) installed). 

If you want to use the shiny app interface, you will also need to install the `shiny` and `shinycssloaders` libraries

# Using the Shiny app

To use the Shiny app, just run the following code in R:
```
library(shiny)
runGitHub('jhelvy/stanTuner')
```

# Example without the Shiny app

(see the `examples.R` file for more examples)

Let's say I want to find the parameters of a normal distribution such that P[x < -2.0] ~ 0.01 and P[x > 2.0] ~ 0.01. That is, I want a normal distribution where 98% of the probability density is between (-2, 2).

First, load the `rstan` library, tweak some settings, and source the `functions.R` file (all functions are loaded in a new environment called `funcs`):

```{r warning=FALSE, message=FALSE}
# Load libraries
library(shiny)
library(rstan)

# Stan settings
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

# Load the functions
funcs <- new.env()
source('functions.R', local=funcs)
```

Then use the `targets` argument to set the desired tail properties:
```{r}
targets = list(
    bound_L = -2,   # LOWER quantile boundary
    bound_U = 2,    # UPPER quantile boundary
    dens_L  = 0.01, # Target density below LOWER quantile boundary
    dens_U  = 0.01) # Target density above UPPER quantile boundary
```

Then use the `tuneParams` function to find the parameters:
```{r message=FALSE}
results = funcs$tuneParams(distribution='normal', targets)
```

View the resulting parameters and verify that the quantiles of 10,000 draws from the resulting distribution match your criteria:
```{r}
results$params
results$quantiles
```

Finally, view a histogram of the resulting distribution:
```{r}
results$histogram
```

# Explanation of the backend
The meat of this app is in the `functions.R` code. The main function is the `tuneParams()` function. After the user defines the `distribution` and the `targets` for the quantiles and density, `tuneParams()` calls the `generateStanCode()` function to generate the Stan code for the model, which is written to the file `model.stan`. This file is always overwritten every time the `generateStanCode()` function is called. It then calls the `stan` function to fit the model. Finally, once the model is fit, it calls the `summarizeResults()` function to extract the results of the fit model. The output of `tuneParams()` is a list with the following values:

* `params` : The fit parameters
* `draws` : 10,000 draws from the resulting distribution using the fit parameters
* `quantiles` : The quantiles of the draws at the desired upper and lower quantile boundaries 
* `histogram` : A histogram of the draws

# Author and License
* Author: John Paul Helveston (www.jhelvy.com)
* Date First Written: Tuesday, April 30, 2019
* License: GPL-3